% This file is part of GUFI, which is part of MarFS, which is released
% under the BSD license.
%
%
% Copyright (c) 2017, Los Alamos National Security (LANS), LLC
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without modification,
% are permitted provided that the following conditions are met:
%
% 1. Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
%
% 2. Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation and/or
% other materials provided with the distribution.
%
% 3. Neither the name of the copyright holder nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND
% ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
% WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED.
% IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
% INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING,
% BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
% DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE
% OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF
% ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
%
%
% From Los Alamos National Security, LLC:
% LA-CC-15-039
%
% Copyright (c) 2017, Los Alamos National Security, LLC All rights reserved.
% Copyright 2017. Los Alamos National Security, LLC. This software was produced
% under U.S. Government contract DE-AC52-06NA25396 for Los Alamos National
% Laboratory (LANL), which is operated by Los Alamos National Security, LLC for
% the U.S. Department of Energy. The U.S. Government has rights to use,
% reproduce, and distribute this software.  NEITHER THE GOVERNMENT NOR LOS
% ALAMOS NATIONAL SECURITY, LLC MAKES ANY WARRANTY, EXPRESS OR IMPLIED, OR
% ASSUMES ANY LIABILITY FOR THE USE OF THIS SOFTWARE.  If software is
% modified to produce derivative works, such modified software should be
% clearly marked, so as not to confuse it with the version available from
% LANL.
%
% THIS SOFTWARE IS PROVIDED BY LOS ALAMOS NATIONAL SECURITY, LLC AND CONTRIBUTORS
% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO,
% THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL LOS ALAMOS NATIONAL SECURITY, LLC OR
% CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL,
% EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT
% OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING
% IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY
% OF SUCH DAMAGE.



\section{\gufistats}
\gufistats is used to analyze a tree and retrieve statistics from an
index. \gufistats does not have an analogous standard utility.

\subsection{Flags}
\gufistats is called with at minimum one positional argument
specifying the statistic desired. Each statistic has a default
version, Most also have a more in-depth version that is specified by
their category: recursive, cumulative, or other.

\begin{table}[h!]
  \centering
  \begin{tabularx}{\textwidth}{| l | X |}
    \hline
    Optional Flags & Functionality \\
    \hline
    -{}-help & displays help menu \\
    \hline
    -v, -{}-version & display program's version number and exits \\
    \hline
    -r, -{}-recursive & run command recursively \\
    \hline
    -c, -{}-cumulative & return cumulative values \\
    \hline
    -{}-order order & sort output (if applicable) \\
    \hline
    -{}-num-results n & first n results \\
    \hline
    -{}-uid u, -{}-user u & restrict to user \\
    \hline
    -{}-reftime REFTIME & time used to compare timestamps with \\
    \hline
    -{}-delim c & Delimeter separating output
    columns \\
    \hline
    -{}-skip-file filename & Name of file containing
    directory basenames to skip \\
    \hline
    -V, -{}-verbose & Show the \gufiquery being executed \\
    \hline
  \end{tabularx}
\end{table}

\subsubsection{Recursive}
These statistics will run a computation on a single
directory. Specifying \texttt{-r} or \texttt{-{}-recursive} will
recursively descend the starting directory and return each
subdirectory's statistic.
\\
\begin{tabularx}{\textwidth}{| l | X |}
  \hline
  Statistic & Description \\
  \hline
  depth & Get the depths of the provided directory relative to the root
  directory \\
  \hline
  filesize & Get the size the files in the immediate directory \\
  \hline
  filecount & Get the number of files in the immediate directory \\
  \hline
  linkcount & Get the number of links in the immediate directory \\
  \hline
  dircount & Get the number of directories in the immediate directory \\
  \hline
  leaf-dirs & Get the leaf directories immediately under the current
  directory \\
  \hline
  leaf-depth & Get the depth of the leaf directories immediately under
  the current directory \\
  \hline
  leaf-files & Get number of files in the leaf directories immediately
  under the current directory \\
  \hline
  leaf-links & Get number of links in the leaf directories immediately
  under the current directory \\
  \hline
\end{tabularx}

\subsubsection{Cumulative}
These statistics will run a computation on an entire subtree. By
default, the results will be grouped together by UID. Specifying
\texttt{-c} or \texttt{-{}-cumulative} will combine all of the results
into a single sum.
\\
\begin{tabularx}{\textwidth}{| l | X |}
  \hline
  Statistic & Description \\
  \hline
  total-filesize & Get the total size taken up by the entire directory
  \\
  \hline
  total-filecount & Get the total number of files under the directory \\
  \hline
  total-linkcount & Get the total number of links under the directory \\
  \hline
  total-dircount & Get the total number of directories under the
  directory \\
  \hline
  total-leaf-files & Get total number of files in the leaf directories
  under this directory \\
  \hline
  total-leaf-links & Get total number of links in the leaf directories
  under this directory \\
  \hline
  files-per-level & Get counts of how many files are in each level of
  the tree \\
  \hline
  links-per-level & Get counts of how many links are in each level of
  the tree \\
  \hline
  dirs-per-level & Get counts of how many dirs are in each level of the
  tree \\
  \hline
  average-leaf-files & Get average number of leaf files under the
  provided directory \\
  \hline
  average-leaf-links & Get average number of leaf links under the
  provided directory \\
  \hline
\end{tabularx}

\subsubsection{Both}
These statistics can be called with both \texttt{-{}-recursive} or
\texttt{-{}-cumulative}. The statistics that have an associated
\texttt{-hist} version call user defined functions and return single
formatted records (see \texttt{histogram.h}) instead of per-bucket
records.
\\
\begin{longtable}{|l|p{0.55\linewidth}|}
  \hline
  extensions & Generate histogram of file extensions \\
  extensions-hist & Return a single formatted record \\
  \hline
  filesize-log2-bins & Generate histogram of filesizes with bins
  [2\textsuperscript{n}, 2\textsuperscript{n+1}) \\
  filesize-log2-bins-hist & \\
  \hline
  filesize-log1024-bins & Generate histogram of filesizes with bins
  [1024\textsuperscript{n}, 1024\textsuperscript{n+1}) \\
  filesize-log1024-bins-hist & \\
  \hline
  dirfilecount-log2-bins & Generate histogram of files per directory with bins
  [2\textsuperscript{n}, 2\textsuperscript{n+1}) \\
  dirfilecount-log2-bins-hist & \\
  \hline
  dirfilecount-log1024-bins & Generate histogram of files per directory with bins
  [1024\textsuperscript{n}, 1024\textsuperscript{n+1}) \\
  dirfilecount-log1024-bins-hist & \\
  \hline
  file-age-atime & Generate histogram of distance from atime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  file-age-atime-hist & \\
  \hline
  file-age-mtime & Generate histogram of distance from mtime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  file-age-mtime-hist & \\
  \hline
  file-age-ctime & Generate histogram of distance from ctime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  file-age-ctime-hist & \\
  \hline
  dir-age-atime & Generate histogram of distance from atime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  dir-age-atime-hist & \\
  \hline
  dir-age-mtime & Generate histogram of distance from mtime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  dir-age-mtime-hist & \\
  \hline
  dir-age-ctime & Generate histogram of distance from ctime. Use
  \texttt{-{}-reftime} to change the timestamp being compared with. \\
  dir-age-ctime-hist & \\
  \hline
\end{longtable}

\subsubsection{Other}
Statistics under the ``other'' category do not have extra flags
associated with them.
\\
\begin{tabularx}{\textwidth}{| l | X |}
  \hline
  Statistic & Description \\
  \hline
  median-leaf-files & Get median number of leaf files under the provided
  directory \\
  \hline
  duplicate-names & Find files with matching names and sizes \\
  \hline
\end{tabularx}
