% This file is part of GUFI, which is part of MarFS, which is released
% under the BSD license.
%
%
% Copyright (c) 2017, Los Alamos National Security (LANS), LLC
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without modification,
% are permitted provided that the following conditions are met:
%
% 1. Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
%
% 2. Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation and/or
% other materials provided with the distribution.
%
% 3. Neither the name of the copyright holder nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND
% ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
% WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED.
% IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
% INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING,
% BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
% DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE
% OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF
% ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
%
%
% From Los Alamos National Security, LLC:
% LA-CC-15-039
%
% Copyright (c) 2017, Los Alamos National Security, LLC All rights reserved.
% Copyright 2017. Los Alamos National Security, LLC. This software was produced
% under U.S. Government contract DE-AC52-06NA25396 for Los Alamos National
% Laboratory (LANL), which is operated by Los Alamos National Security, LLC for
% the U.S. Department of Energy. The U.S. Government has rights to use,
% reproduce, and distribute this software.  NEITHER THE GOVERNMENT NOR LOS
% ALAMOS NATIONAL SECURITY, LLC MAKES ANY WARRANTY, EXPRESS OR IMPLIED, OR
% ASSUMES ANY LIABILITY FOR THE USE OF THIS SOFTWARE.  If software is
% modified to produce derivative works, such modified software should be
% clearly marked, so as not to confuse it with the version available from
% LANL.
%
% THIS SOFTWARE IS PROVIDED BY LOS ALAMOS NATIONAL SECURITY, LLC AND CONTRIBUTORS
% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO,
% THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL LOS ALAMOS NATIONAL SECURITY, LLC OR
% CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL,
% EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT
% OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING
% IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY
% OF SUCH DAMAGE.



\section{Indexing}
The first step to using GUFI is indexing a source filesystem.

An index created by GUFI retains the shape of the source filesystem:
directories that exist in the source filesystem also exist in the
index. If an administrator created the index, the directories will
also have the same access permissions, uid, and gid.

Indexes will not contain any of the files that the source filesystem
contained. Instead, all metadata extracted from a single directory of
the source filesystem will be placed into a single database file,
called db.db, in the cooresponding directory in the index. Each
database file will be created with a fixed schema that includes the
tables listed in Sections \ref{sec:schema} and
\ref{sec:xattr_schema}. Additional database files may be created if
extended attributes are extracted (see Section~\ref{sec:xattrs}).

Index processing (creation) occurs on a per-directory basis, and thus
is highly parallelizable.

\subsection{Directly Indexing a Filesystem}
\input{sections/gufi_dir2index}
\subsection{Indirectly Indexing a Filesystem}
\input{sections/tracefile}
\input{sections/gufi_dir2trace}
\input{sections/gufi_trace2index}

\subsection{Index Placement}
The output (final positional) argument to \gufidirindex and
\gufitraceindex does not point to the directory where the index
starts. Rather, it points to the parent directory of where the index
starts\footnote{An empty db.db file is also placed here.}. This was
done in order to allow for users to pick the name of their GUFI search
space while leaving the index's top-most directory with the same name
as the source tree's top-most directory.

Placing multiple indexes under the same prefix is as simple as
maintaining the same output argument with different inputs
paths/traces.

Placing an index into the source tree is possible by pointing
\gufidirindex to the parent of the source tree.

Pointing the output of \gufidirindex to a location under the source
tree may result in the index creation never completing due to index
directories being created below the current tree walk location as the
source tree is being walked, causing the index directories to be
indexed, creating more index directories to be walked ad infinitum.

\input{sections/xattrs}
\input{sections/external_databases}

\subsection{Location}
GUFI is expected to be used to query the indexes of many filesystems
all at once. However, the source filesystems are not expected to be
accessible from each other. In order for the indexes from many
disconnected systems to be queried at once, they should all be built
under a common directory on a single machine.

\subsubsection{Permissions}
If a GUFI tree is located on a different machine than the source
tree, the users and groups are probably not available on the machine
with the index on it. \texttt{/etc/passwd} should be populated with
the entries from the source machine and modified so that they do not
have a home directory, and are not allowed to log in
(\texttt{/sbin/nologin}). Similarly, \texttt{/etc/group} should be
updated with the source machine's groups and modified to remove any
unnecessary information.
