% This file is part of GUFI, which is part of MarFS, which is released
% under the BSD license.
%
%
% Copyright (c) 2017, Los Alamos National Security (LANS), LLC
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without modification,
% are permitted provided that the following conditions are met:
%
% 1. Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
%
% 2. Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation and/or
% other materials provided with the distribution.
%
% 3. Neither the name of the copyright holder nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND
% ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
% WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED.
% IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
% INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING,
% BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
% DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE
% OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF
% ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
%
%
% From Los Alamos National Security, LLC:
% LA-CC-15-039
%
% Copyright (c) 2017, Los Alamos National Security, LLC All rights reserved.
% Copyright 2017. Los Alamos National Security, LLC. This software was produced
% under U.S. Government contract DE-AC52-06NA25396 for Los Alamos National
% Laboratory (LANL), which is operated by Los Alamos National Security, LLC for
% the U.S. Department of Energy. The U.S. Government has rights to use,
% reproduce, and distribute this software.  NEITHER THE GOVERNMENT NOR LOS
% ALAMOS NATIONAL SECURITY, LLC MAKES ANY WARRANTY, EXPRESS OR IMPLIED, OR
% ASSUMES ANY LIABILITY FOR THE USE OF THIS SOFTWARE.  If software is
% modified to produce derivative works, such modified software should be
% clearly marked, so as not to confuse it with the version available from
% LANL.
%
% THIS SOFTWARE IS PROVIDED BY LOS ALAMOS NATIONAL SECURITY, LLC AND CONTRIBUTORS
% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO,
% THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL LOS ALAMOS NATIONAL SECURITY, LLC OR
% CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL,
% EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT
% OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING
% IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY
% OF SUCH DAMAGE.



\subsection{\qptp}
Thread pools are generally written with a single work thread (with a
single lock) from which to threads pull from. This does not work past
a handful of threads. GUFI can run on hundreds of threads, and thus
required a more performant thread pool. The solution to this was
\qptp.

\qptp is named so because originally, there was one work queue (and
lock) maintained for each thread in the thread pool. Now, there are
multiple work queues maintained for each thread: work is pushed to the
waiting work queue and the claimed work queue contains the work that
the thread claimed for processing.

\subsubsection{Adding Work}
By default, work is added to threads in a round robin fashion in order
to distribute work evenly and to attempt to prevent, or at least
reduce the amount of, contention experienced by any one work queue.

This function can be changed during initialization.

\subsubsection{Processing Queued Work}
Because work items are enqueued in parallel while work items are
processed, popping off work items one at a time results in one lock
per removal that might experience contention while the queue is being
modified. In \qptp, when a thread claims (pops off) work for
processing, {\bf all} work items in the waiting work queue are moved
to the claimed work queue, resulting in the removal of multiple work
items with only a single lock that might experience contention. All
claimed work is processed before the thread returns to the work queue
to find more work.

If a thread discovers that it does not have work items in its waiting
work queue but the thread pool still has outstanding in-memory work,
it will search for more work in the waiting queue of other threads. If
work items are found, the thread will steal at least one item, causing
the stolen work items to experience less latency between enqueuing and
processing. If no work can be stolen from the waiting work queues of
all threads, the thread will attempt to take work from the claimed
work queue. This is done so that long running jobs do not prevent
claimed work from being processed. This also makes \qptp behave more
like thread pools that pull one work item off at a time and thus
always processes work when work is available.

If the queue size limit was set before the thread pool was started and
a thread's work item count is greater than or equal to the queue size
limit, new work items are swapped to a filesystem instead of cached in
memory. These work items are not counted as in-memory outstanding
work, and are instead processed after all in-memory work items have
been processed, in \texttt{QPTPool\_wait}. This should be considered a
last-ditch attempt to prevent OOM-ing, and should not occur too often.

\subsubsection{Usage}
\begin{enumerate}
\item Create a thread pool:

  \texttt{QPTPool\_ctx\_t *ctx = QPTPool\_init(nthreads, args);}

  \texttt{nthreads} sets the number of threads in this thread
  pool.

  The \texttt{args} argument will be accessible by all threads
  processing work items. Use
  \texttt{QPTPool\_get\_args\_internal(ctx)} to retreive it from
  within a thread.

  The \texttt{ctx} returned is an array of thread contexts but is
  opaque to the caller. Enqueuing new work from within a user function
  uses the context for that particular thread to figure out where to
  put the new work item. In order to not have a different function for
  enqueuing work from outside of the \texttt{QPTPool}, say from
  \texttt{main()}, the returned array also acts as the context of
  thread 0, taking advantage of the fact that \texttt{ctx == \&ctx[0]}.

\item Setting Properties:

  \texttt{QPTPool\_init} is intentionally kept simple and uses default
  values for some features. These values may be set using
  \texttt{QPTPool\_init\_with\_props} or modified using the
  \texttt{QPTPool\_set\_*} functions before \texttt{QPTPool\_start} is
  called.

  By default, \qptp will push new work items in a round robin
  fashion. This can be changed to a custom function with
  \\\texttt{QPTPool\_set\_next}. This function is set at the context
  level instead of at \texttt{QPTPool\_enqueue} in order to not
  require a branch to figure out whether or not the provided function
  pointer is valid.

  \texttt{QPTPool\_set\_queue\_limit} causes work items to be swapped
  to a filesystem.

  \texttt{QPTPool\_set\_swap\_prefix} sets the path where swap data is
  written. Note that this prefix is prefixed to the path string
  directly, without a path separator. Additionally, the parent
  directory of the prefix must already exist - the path is not created
  if it does not exist.

  \texttt{QPTPool\_set\_steal} sets the numerator and denominator of
  the multiplier used when work items are being stolen from other
  threads. If the queue being inspected has at least one work item on
  it, \texttt{max(queue.size * numerator / denominator, 1)} work items
  are taken from the front of the queue.

\item Getting Properties:

  Properties may be extracted from the context using the
  \texttt{QPTPool\_get\_*} functions.

  Note that two functions, \texttt{QPTPool\_get\_id} and
  \texttt{QPTPool\_get\_args\_internal}, are intended to be used
  within the function processing a work item, and thus the context is
  assumed to be valid, and is not checked for \texttt{NULL}s.

\item Start the thread pool:

  \texttt{QPTPool\_start(ctx);}

\item Add work:

  \begin{itemize}
    \item Each of the following functions take at least the following
      arguments:
      \begin{itemize}
        \item \texttt{ctx}: The context returned by \texttt{QPTPool}
          initialization.
        \item \texttt{function}: The function that will be used to
          process the associated work item. \texttt{int
            (*)(QPTPool\_ctx\_t *ctx, void *data)} typedef-ed to
          \texttt{QPTPool\_f},.
        \item \texttt{work}: A pointer to the data that will be
          processed by this thread. This is renamed to \texttt{data}
          when passed to the function.
      \end{itemize}

    \item \texttt{QPTPool\_enqueue(ctx, function, work);}

    The thread that will receive the new work item is not
    \texttt{ctx}. Rather, \texttt{ctx} is used to figure out where
    the new work item is enqueued.

    Work enqueued with \texttt{QPTPool\_enqueue} will not be swapped.

    \item \texttt{QPTPool\_enqueue\_swappable(ctx, func, work, serialize, deserialize);}

    Work passed into \texttt{QPTPool\_enqueue\_swappable} may end up
    in the wait queue or be swapped to a filesystem.

    The thread that will receive the new work item is not
    \texttt{ctx}. Rather, \texttt{ctx} is used to figure out where
    the new work item is enqueued.

    \item \texttt{QPTPool\_enqueue\_here(ctx, id, queue, func, work, serialize, deserialize);}

    If the queue argument is set to \texttt{QPTPool\_enqueue\_WAIT},
    the work will be explicitly placed into the selected queue at
    thread \texttt{id} instead of using the next thread selection
    function. The queue limit is ignored.

    If the queue argument is set to \texttt{QPTPool\_enqueue\_SWAP},
    the work will be written to the selected thread's swap space. The
    \texttt{serialize} and \texttt{deserialize} functions are used to
    write and read the work data to and from the filesystem.

  \end{itemize}

\item Wait for work to be completed (optional):

  \begin{itemize}
    \item \texttt{QPTPool\_wait\_mem\_lte(ctx, count);}

      This function waits until the amount of in-memory work to be
      done is less than or equal to \texttt{count}.

  \item \texttt{QPTPool\_wait\_mem(ctx);}

      This function waits until all in-memory work to be done. This
      function is just \texttt{QPTPool\_wait\_mem\_lte(ctx, 0);}

    \item \texttt{QPTPool\_wait(ctx);}

      This function waits until all in-memory and swapped work is done.
  \end{itemize}

\item Join threads after all work has been completed:

  \texttt{QPTPool\_stop(ctx);}

  This function exists to allow for the collection of statistics
  before the context is destroyed. Note that \texttt{QPTPool\_wait} is
  called by this function, and thus does not have to be explicitly
  called before calling this function.

\item Destroy the pool context:

  \texttt{QPTPool\_destroy(ctx);}
\end{enumerate}
