% This file is part of GUFI, which is part of MarFS, which is released
% under the BSD license.
%
%
% Copyright (c) 2017, Los Alamos National Security (LANS), LLC
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without modification,
% are permitted provided that the following conditions are met:
%
% 1. Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
%
% 2. Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation and/or
% other materials provided with the distribution.
%
% 3. Neither the name of the copyright holder nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" AND
% ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED
% WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED.
% IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT,
% INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING,
% BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
% DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE
% OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF
% ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
%
%
% From Los Alamos National Security, LLC:
% LA-CC-15-039
%
% Copyright (c) 2017, Los Alamos National Security, LLC All rights reserved.
% Copyright 2017. Los Alamos National Security, LLC. This software was produced
% under U.S. Government contract DE-AC52-06NA25396 for Los Alamos National
% Laboratory (LANL), which is operated by Los Alamos National Security, LLC for
% the U.S. Department of Energy. The U.S. Government has rights to use,
% reproduce, and distribute this software.  NEITHER THE GOVERNMENT NOR LOS
% ALAMOS NATIONAL SECURITY, LLC MAKES ANY WARRANTY, EXPRESS OR IMPLIED, OR
% ASSUMES ANY LIABILITY FOR THE USE OF THIS SOFTWARE.  If software is
% modified to produce derivative works, such modified software should be
% clearly marked, so as not to confuse it with the version available from
% LANL.
%
% THIS SOFTWARE IS PROVIDED BY LOS ALAMOS NATIONAL SECURITY, LLC AND CONTRIBUTORS
% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO,
% THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL LOS ALAMOS NATIONAL SECURITY, LLC OR
% CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL,
% EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT
% OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING
% IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY
% OF SUCH DAMAGE.



\subsection{Generate \treesummary Tables}
\label{sec:treesummary}
The \treesummary table is an optional table that is placed into
db.db. It contains a summary of the entire subtree starting at current
directory using minimums, maximums, and totals of numerical values.

When a query is provided to \gufiquery \texttt{-T}, the \treesummary
table is queried first. Because \treesummary tables do not necessarily
exist, \gufiquery first checks for the existence of the \treesummary
table in the directory being processed before performing the
\texttt{-T} query. If the \texttt{-T} query returns no results, the
entire subtree will be skipped.

\subsubsection{\gufitreesummary}
Starting from a directory provided in the command line,
\gufitreesummary recursively traverses to the bottom of the tree,
collecting data from the \summary table of each child directory
database. If a \treesummary table is discovered in a subdirectory,
descent down the tree is stopped as the \treesummary table contains
all of the information about that directory as well as all
subdirectories underneath it. Once all of the data has been collected,
it is summarized and placed into the \treesummary table of the
starting directory.

Generating \treesummary tables for all directories using this top-down
approach will take a long time due to the repeated traversals across
the same directories. Because of this, \gufitreesummary generates the
\treesummary table for the provided directory only.

If generating \treesummary tables using \gufitreesummary, the tables
should be generated at optimal points within the index. For example,
if the index is on a home directory, it may be useful to generate
\treesummary tables at each user's home directory.
\\\\
Example Call:
\\\\
\indent \gufitreesummary \indexroot

\subsubsection{\gufitreesummaryall}
\gufitreesummaryall generates \treesummary tables for all
directories. This is done by walking to the bottom of the tree and
generating \treesummary tables while walking back up, which only
occurs after all subdirectories have had their \treesummary tables
generated. Leaf directories, by definition, do not have
subdirectories, and further traversal down is unnecessary and
impossible. Their \treesummary tables are thus duplicates of their
\summary tables, providing the base case for the walk back up the
tree. Directories above the leaves can then use the \treesummary data
found in their immediate subdirectories, which are 1) guaranteed to
exist and 2) guaranteed to summarize the entire subdirectory's
subtree, to generate their own \treesummary tables.
\\\\
Example Call:
\\\\
\indent \gufitreesummaryall \indexroot

\subsubsection{\gufirollup}
Just as with \gufitreesummaryall, rolling up a tree involves walking
to the bottom of the tree and working upwards. This allows for
\treesummary generation to be performed automatically during the roll
up operation, resulting in \treesummary tables being generated for all
directories whether or not they were rolled up.
